Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences

### 2019-02-08 draft.tex
more mess into the tex

\title{Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences}
%\author{}
\author{Olivier Cailloux \and Stefano Moretti \and Beatrice Napolitano \and Paolo Viappiani} %\affil{LIP6\\\email{paolo.viappiani@lip6.fr}
%\author{Olivier Cailloux} 
%\author{Stefano Moretti}%\affil{LAMSADE}
%\author{Beatrice Napolitano}\affil{LAMSADE}

@@

\begin{itemize}
\item General motivation: need to deal with partially specified preferences and as well partially specified voting rules
\item Focus on positional scoring rules
\item Robust decision criterion: minimax regret
\item Incremental elicitation of preferences can help mitigating the cognitive and communication costs
\item Novelty of the proposed method
\item Structure of the paper
\item References: \cite{Cailloux2014}, \cite{Llamazares2013}, \cite{Konczak05}, \cite{Kalech2011}, \cite{Pini2009}, \cite{Xia2008}
\end{itemize}

@@

The traditional approach to social choice assumes that both the social function and the full preference orderings of the voters are expressed beforehand. 
As noted by several authors requiring voters to express full preference orderings can be onerous (especially for large sets of alternatives)

% Given partial information about voter preferences, how does one choose a suitable approximate winner? 
Scoring rules are a particularly common methods used to aggregate rankings and determine  a winner.
The score can be seen as a utility for the society...

%and often captures more information than is needed to determine the winner
We develop methods for computing the minimax- optimal alternative for scoring rules with partial preference information and partial information about the weights.

%which queries to ask of which voters. 
We discuss  heuristics that allow us to determine queries that quickly allow one to reduce minimax regret.

 Our experimental results on randomly generated and real world data sets show...

In this paper we propose an interactive adaptive protocol for eliciting both the voters preferences and the voting rule.
At each step we need to decide whether we want to ask a question to the committee or to one of the agents (and to which agent in particular).
%We want to design a questioning strategy to minimize the worst regret in the long run. A questioning strategy indicates, given a partial profile and a set of feasible weights, which question to ask next. A question is either a query to the voter $i$ about her preference relation $\succ_i$ or a query to the committee about the feasible weights.

@@

\subsection{Related works}

\paragraph{Uncertain scoring rules}
A number of works have dealt with the problem of reasoning with incompletely specified aggregation functions.
In particular, when considering positional scoring rules, it is possible to derive dominance relations (akin to stochastic dominance) that allow to eliminate some alternatives since they will be less preferred than another one for any instantiation of the weights \cite{Stein1994}.
More recently the characterization of methods for aggregating the uncertainty over the scoring vectors has been studied in \cite{Viappiani2018}.
We also note that the elicitation methods based on minimax regret described in \cite{Boutilier2006} (and in many other recent papers), while not specifically targeted to scoring rules, can be easily adapted (from a technical point of view) to elicit the scoring vector from a committee.

@@

A social choice function $f : V \rightarrow 2^A$ associates a profile with a set of winners.
Among the many possible social choice functions, we consider {\em positional scoring rules}, which attach weights to positions; an alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
= \sum_{i=1}^{m} \alpha_i w_i . \label{eq:srule}
\end{align}
where $\alpha_i$ is the number of times that alternative $x$ was ranked in the $i$-th position.
The vector $(w_1,\ldots,w_m)$ is called the scoring vector. We assume that the weights constitute a monotonic sequence: $w_{1} \geq w_{2} \geq \ldots \geq w_{m}$.

@@

%The regret-minimizing alternative may not be a possible winner for some voting rules.
The weights of a scoring rule can model different preferences of the committee. 
For instance, the weights can control the inclination to favour ``extreme'' alternatives (often at either the top or the bottom of the rankings) at the expenses of ``moderate'' alternatives (that are more consistently in the middle part of the rankings).

However in general it can be difficult to set the weights in an appropriate way.

@@

We propose to use the minimax regret as a decision criterion to determine a winner.
Minimax regret [ref] has been used for making decisions under utility uncertainty and for preference elicitation [ref].
Lu and Boutilier \cite{Lu2011} have adopted minimax regret ;

@@

A scoring rule associates each alternative to a score that is given by the sum of points obtained for each voter.

Scoring rules are additively decomposable.
Let $s(x; v_j,\w)=w_{v_j(x)}$ be the number of points that $x$ obtains in the ranking $v_j$ (see Equation \ref{eq:srule}), that is the weight of position $v_j(x)$.
Obviously we have 
\[ s(x; \profile, \w) = \sum_{j=1}^n s(x; v_j,\w). \]

@@

We compute pairwise maximum regret by considering binary variables to represent optimization choices related to where to position the alternatives.
We also have numerical variables to represent the weights of the scoring rule.

Since the objective is to maximize pairwise regret...

 The objective function can be written as:
 \[ \max \sum_{l=1}^n I^j_l w_l -  I^j_l w_l \]

 We replace the multiplicative terms by new variables..
\[ V_{l}^{j} \sim I^j_l w_l  \]

To impose the desired behavior:
\begin{align*} 
V_{l}^{j} \leq I^j \\
V_{l}^{j} \leq w_l
\end{align*}
The expression $I^j_l w_l$ evaluates to $w_l$ if $I^j_l = 1$, and to $0$ otherwise.

We write $w \in W$ as a shourtcut to represent the constraints that the weights are chosen to be in the feasible set.
%Variable $I^j$ associated with the constraint $0 \leq I \leq |U^j|$.
 Finally, the following is the optimization problem that compute the value $\PMR$.
 The variables are $w$, $V$, $I$,...
\begin{align}
\max & \sum_{l=1}^m  V_{l}^{j,+} - V_{l}^{j,-}\\
\text{ s.t. } & w \in W  \\
& \sum_{l} I^{j}_{l} = 1 & \forall j \\
& V_{l}^{j,+} \leq I^j  & \forall l,j \\
& V_{l}^{j,+} \leq w_l & \forall l,j \\
& V_{l}^{j,-} \geq w + I^j - 1 & \forall l,j \\
& V_{l}^{j,-} \geq 0 & \forall l,j\\
\end{align} 
The optimization program can be solved by any suitable MILP solver.

@@

%regret-based incremental elicitation method
\begin{itemize}
	\item Minimax regret can be used to guide the elicitation process
	\item As termination condition of elicitation, we can check whether minimax regret is lower than a threshold. If optimality is derived, we can perform elicitation until minimax regret drops to zero.
	\item At each step, the system need to decide whether to ask a question to one of the agents or ask a question to the chair about the voting rule
	\item Type of questions that we can ask to the chair:
	bound queries (Is this alternative among your top-k most preferred items?), comparison queries (do you prefer alternative x or alternative y?).
\end{itemize}

@@

\paragraph{Query methods}

%{\em Some ideas: decompose the regret into two components, one due to $\w$ and one due to $\pprofile$, and ask a question to the chair / or to one of the agents depending on which is highest}

@@

@@

% We can stop elicitation when minimax regret is sufficiently small.
We evaluate the proposed method using the following simulated protocol.

We test the elicitation protocols in randomly generated datasets and real datasets (sushi and iris).

@@

\section{Conclusions}  \label{sec:conclusions}
%  Possible future works: manipulation,

This is an awesome paper, it should be accepted right away and win a prize.

@@

\subsection{Profile completion, again}
 %{\bf \em TODO: still have to find out exactly how to do it}

 We introduce a set integer variables $I^i$, one for each position $i$.
 Let $U^j$ be, for the agents in $A^-$, the alternatives that, in the partial order $\succ^p_j$ of some agent $j \in A^-$, are incomparable with $x$ and $y$.
 \[ \sum_{l=1}^n I^j_l w_l \]
 Variable $I^j$ associated with the constraint $0 \leq I \leq |U^j|$.

 We provide the following MIP (mixed integer linear program).

% Use linearization techniques to handle the multiplications...

%\section{Even more fun}
%We can consider different types of questions: asking to compare a pair of alternatives, asking about top-k alternatives.
%

>>>>

\title{Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences}
%\author{}
\author{Olivier Cailloux  \and Beatrice Napolitano \and Paolo Viappiani} %\affil{LIP6\\\email{paolo.viappiani@lip6.fr}
%\author{Olivier Cailloux} 
%\author{Stefano Moretti}%\affil{LAMSADE}
%\author{Beatrice Napolitano}\affil{LAMSADE}

@@

\begin{itemize}
\item General motivation: need to deal with partially specified preferences and as well partially specified voting rules
\item Focus on positional scoring rules
\item Robust decision criterion: minimax regret
\item Incremental elicitation of preferences can help mitigating the cognitive and communication costs
\item Novelty of the proposed method
\item References: \cite{Cailloux2014}, \cite{Llamazares2013}, \cite{Konczak05}, \cite{Kalech2011}, \cite{Pini2009}, relation to necessary and possible winners \cite{Xia2008}
\end{itemize}

@@

The traditional approach to social choice assumes that both the social function and the full preference orderings of the voters are expressed beforehand. 
As noted by several authors requiring voters to express full preference orderings can be onerous (especially for 
large sets of alternatives).
Acquiring the necessary preference information may require a high cognitive and communication cost.

 In many real situations, however, it may be necessary to reason with partial preferences, as some prefer- ences are not available and too expensive to obtain (with respect to a cognitive or economic cost). 
This observation has motivated a number of recent works on

In this paper we relax the traditional assumptions.
We develop methods for computing the minimax- optimal alternative for scoring rules with partial preference information and partial information about the weights.

We an adaptive utility elicitation

We focus on positional scoring rules, that  are a particularly common methods used to aggregate rankings and determine  a winner.
We also address the problem of elicitation, providing incremental elicitation methods to acquire relevant preference information.%The score can be seen as a utility for the society.
%and often captures more information than is needed to determine the winner
We discuss  heuristics that allow us to determine queries that quickly allow to reduce minimax regret.

%In this paper we propose an interactive adaptive protocol for eliciting both the voters preferences and the voting rule.
%At each step we need to decide whether we want to ask a question to the committee or to one of the agents (and to which agent in particular).

%We want to design a questioning strategy to minimize the worst regret in the long run. A questioning strategy indicates, given a partial profile and a set of feasible weights, which question to ask next. A question is either a query to the voter $i$ about her preference relation $\succ_i$ or a query to the committee about the feasible weights.

@@

%\subsection{Related works}
\paragraph{Uncertain scoring rules}
A number of works have dealt with the problem of reasoning with incompletely specified aggregation functions.
In particular, when considering positional scoring rules, it is possible to derive dominance relations (akin to stochastic dominance) that allow to eliminate some alternatives since they will be less preferred than another one for any instantiation of the weights \cite{Stein1994}.
More recently the characterization of methods for aggregating the uncertainty over the scoring vectors has been studied in \cite{Viappiani2018}.
We also note that the elicitation methods based on minimax regret described in \cite{Boutilier2006} (and in many other recent papers), while not specifically targeted to scoring rules, can be easily adapted (from a technical point of view) to elicit the scoring vector from a committee.

@@

A social choice function $f : V \rightarrow 2^A$ associates a profile with a set of winners.
Among the many possible social choice functions, we consider {\em positional scoring rules}, which attach weights to positions; 
%A scoring rule associates each alternative to a score that is given by the sum of points obtained for each voter.
an alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
= \sum_{i=1}^{m} \alpha_i w_i . \label{eq:srule}
\end{align}
where $\alpha_i$ is the number of times that alternative $x$ was ranked in the $i$-th position.
The vector $(w_1,\ldots,w_m)$ is called the scoring vector. We assume that the weights constitute a monotonic sequence: $w_{1} \geq w_{2} \geq \ldots \geq w_{m}$.

@@

%The regret-minimizing alternative may not be a possible winner for some voting rules.
The weights of a scoring rule can model different preferences of the committee. 
For instance, the weights can control the inclination to favour ``extreme'' alternatives (often at either the top or the bottom of the rankings) at the expenses of ``moderate'' alternatives (that are more consistently in the middle part of the rankings).
However in general it can be difficult to set the weights in an appropriate way.

@@

We propose to use the minimax regret as a decision criterion to determine a winner.
Minimax regret \cite{Boutilier2006} has been used for making decisions under utility uncertainty and for preference elicitation.
%Lu and Boutilier \cite{Lu2011} have adopted minimax regret.

@@

Scoring rules are additively decomposable.
%A scoring rule associates each alternative to a score that is given by the sum of points obtained for each voter.
Let $s(x; v_j,\w)=w_{v_j(x)}$ be the number of points that $x$ obtains in the ranking $v_j$ (see Equation \ref{eq:srule}), that is the weight of position $v_j(x)$.
Obviously we have 
\[ s(x; \profile, \w) = \sum_{j=1}^n s(x; v_j,\w). \]

@@

We compute pairwise maximum regret by considering binary variables to represent optimization choices related to where to position the alternatives.
We also have numerical variables to represent the weights of the scoring rule.
%Since the objective is to maximize pairwise regret...

 The objective function can be written as:
 \[ \max \sum_{l=1}^n B^j_l w_l -  B^j_l w_l \]

We use enconding trick in order to solve the problem.
We replace the multiplicative terms by new variables.
\[ V_{l}^{j} \sim B^j_l w_l  \]

To impose the desired behavior:
\begin{align*} 
V_{l}^{j} \leq B^j \\
V_{l}^{j} \leq w_l
\end{align*}
The expression $I^j_l w_l$ evaluates to $w_l$ if $I^j_l = 1$, and to $0$ otherwise.

We write $w \in W$ as a shourtcut to represent the constraints that the weights are chosen to be in the feasible set.
%Variable $I^j$ associated with the constraint $0 \leq I \leq |U^j|$.
 Finally, the following is the optimization problem that compute the value $\PMR$.
 The variables are $w$, $V$, $B$,...
\begin{align}
Z := \max & \sum_{l=1}^m  V_{l}^{j,+} - V_{l}^{j,-}\\
\text{ s.t. } & w \in W  \\
& \sum_{B=i_{1}}^{i_{2}} I^{j}_{l} = 1 & \forall j \\
& V_{l}^{j,+} \leq B^j_{l}  & \forall l,j \\
& V_{l}^{j,+} \leq w_l & \forall l,j \\
& V_{l}^{j,-} \geq w + B_{l}^j - 1 & \forall l,j \\
& V_{l}^{j,-} \geq 0 & \forall l,j\\
\end{align} 
The optimization program can be solved by any suitable MILP solver.

\begin{claim}
The $\PMR$ is computed using the above optimization problem.
\end{claim}

@@

%regret-based incremental elicitation method
\begin{itemize}
	\item Minimax regret can be used to guide the elicitation process
	\item As termination condition of elicitation, we can check whether minimax regret is lower than a threshold. If we wish optimality, we can perform elicitation until minimax regret drops to zero.
	\item At each step, the system needs to decide whether to ask a question to one of the agents or ask a question to the chair about the voting rule
	\item Type of questions that we can ask to the chair:
	bound queries (Is this alternative among your top-k most preferred items?), comparison queries (do you prefer alternative x or alternative y?).
\end{itemize}

@@

\paragraph{Query types}

Questions to the voters: it is natural to consider comparison queries asking to compare two alternatives

Questions to the committee: we consider questions that refine our knowledge about the weighting vectors $w$.

\paragraph{Query methods}
We consider different strategies to determine the next question to ask given the current information.
%{\em Some ideas: decompose the regret into two components, one due to $\w$ and one due to $\pprofile$, and ask a question to the chair / or to one of the agents depending on which is highest}

@@

% Determine the part of ``regret'' due to uncertainty in $w$
% Determine the part of ``regret' due to uncertainty in $p$

@@

% We can stop elicitation when minimax regret is sufficiently small.
We test the elicitation protocols in randomly generated datasets and real datasets. % (sushi and iris).
We evaluate the proposed method using the following simulated protocol.

@@

\section{Conclusions}  \label{sec:conclusions}
%  Possible future works: manipulation, 
We have proposed the use of minimax regret as a way to determine a winner in a robust way when considering uncertainty on both the voting rule and the voters' preferences.

Our experimental results on randomly generated and real world data sets show that regret-based elicitation is effective and allow to find a near optimal consensus choice in a limited number of steps.
%Our interactive elicitation methods are shown to be effective in simulations.

@@

%\subsection{Profile completion, again}
% %{\bf \em TODO: still have to find out exactly how to do it}
% 
% We introduce a set integer variables $I^i$, one for each position $i$.
% Let $U^j$ be, for the agents in $A^-$, the alternatives that, in the partial order $\succ^p_j$ of some agent $j \in A^-$, are incomparable with $x$ and $y$.
% \[ \sum_{l=1}^n I^j_l w_l \]
% Variable $I^j$ associated with the constraint $0 \leq I \leq |U^j|$.
% 
% We provide the following MIP (mixed integer linear program).
% 
%% Use linearization techniques to handle the multiplications...
%
%
%%\section{Even more fun}
%%We can consider different types of questions: asking to compare a pair of alternatives, asking about top-k alternatives.
%
